Getting the Ball Rolling: Producing a Foundation Text for a Universal Corpus of the World's Languages

### 2014-03-01 docs/acl-2014/paper.tex
More work on paper.

Subsequently, \newcite{brunello2009free} stressed the notion of building free corpora from the web using documents released under Creative Commons licenses. To build better quality WaCky corpora, \newcite{versley2012quality} introduced the notion content-sensitive boilerplate detection in cleaning web corpora.

The initial wave of monolingual WaCky corpora \cite{baroni2008wacky} inspired more domain specific WaCky corpora, such as the Korean Web Learner's Corpus \cite{dickinson2010korean} and the Feed Corpus \cite{minocha2013feed}. Meanwhile, Web-driven corpora compilations also focused on building large  multilingual resources, e.g. Leipzig Corpora Collection (\texttt{LCC}) \cite{biemann2007leipzig}, the COrpora from the Web (\texttt{COW}) project \cite{schaefer2012cow}.

Different from previous WaCs built under the Web as Corpus kool ynitiative (WaCky), we did not perform seed query web-searching or web crawling to achieve a balanced resource. Instead the foundation text for uniWaC was built using dedicated site crawling and/or site-dependent cleaning from four data sources, viz. (i) \emph{Online Database of Interlinear Text} (ODIN), (ii) \emph{Omniglot} website, (iii) \emph{Universal Declaration of Human Rights} (UDHR) and (iv) \emph{Wikipedia} dumps. Our emphasis for the uniWaC is based on based on \textbf{\emph{universality}}; covering as many languages as possible is the first priority.

@@

\newcite{abney2011data} describe the data structure in more detail, distinguishing between \textbf{aligned texts} and \textbf{analysed texts}: the former consists of multiple parallel texts, aligned at the document, sentence, or word level; the latter contains more detailed annotations including parts of speech, morphological information, and syntactic relations. In our work, we have strived to maintain a data structure consistent with their recommendations for aligned texts.

@@

The resulting files were saved as plaintext file with \texttt{UTF-8} encoding. One directory for each data source and one file per language. The parallel data (from ODIN and Omniglot) are saved as tab-delimited files, one line per sentence, one column per language. The monolingual data (from UDHR and Wikipedia) are saved as newline delimited plaintext file, each line denoting a paragraph or sentence and each document separated by an empty line. The files are saved with the following naming convention to allow easy access, the filename contains the data source and its ISO 639-3 code, e.g. \texttt{odin-iii.txt} contains the \emph{ODIN} data for the \emph{Nuosu} language.

@@

The ODIN data is easily accessible in XML format from the online database\footnote{http://odin.linguistlist.org/download}, where data for each language is saved in a separate XML file and the Interlinerized Glossed Texts (IGTs) are encoded in the \texttt{<igt><example>...</example></igt>} tags. Each XML file is saved under a filename that matches the respective language code. While cleaning the data, filenames that does not adhere to ISO 639-3\footnote{http://www-01.sil.org/iso639-3} were excluded in the uniWaC compilation.

@@

a. \quad o lesu mai \\
\indent \qquad 2sg return here \\
\indent \qquad `\emph{You return here.}' \\

\centerline{Figure 1: IGT adhering to Leipzig Glossing Rule.}
\end{minipage}

\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>a. o lesu mai</line>
    <line>2sg return here</line>
    <line>`You return here.'</line>
  </example>
</igit>
\end{lstlisting} 
\centerline{Figure 2: An IGT in ODIN's XML format.}
\end{minipage}
\\ \\
\noindent The IGTs from a web document usually follow the Leipzig Glossing Rules, where the first line is the source language text, the second line contains the word/morphemic equivalence and the third line is an English gloss. From the ODIN XML format, an IGT as of Figure 1 will be represented as in an XML snippet as in Figure 2. Eventually, we need to clean and extract (i) the source text `\emph{o lesu mai}' without the preceding index `\emph{a.}' and (ii) the target language gloss without the quotation marks `\emph{You return here}'.
\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>(69) na-Na-tmi-kwalca-t 
    (Foley 1991)</line>
    <line>3sgA-1sgO-say-rise-PERF
    </line>
    <line>`She woke me up' 
    (by verbal action)</line>
  </example>
</igit>
\end{lstlisting} 
\smallskip
\centerline{Figure 3: Another IGT in ODIN's XML format.}
\end{minipage}
\\ \\
\noindent The primary problem in extracting the source text is a lack of consistency in the IGTs. In the above examples, the sentence is introduced by a letter or number, which needs to be removed; however, the form of such indexiing elements varies. In addition, the source line in Figure 3 includes a citation for the example, which introduces noise to the corpus.

Lewis and Xia (2010) noted that it is not uncommon for IGTs found in documents to deviate from a single convention; when compiling the ODIN data, they attempted a regex-based approached and reported 59\% F-score. Similar to their regex approaches for extracting IGT, we cleaned the source line with regexes and kept the ODIN data with two levels of `cleanliness':
\\ \\
\begin{minipage}{\columnwidth}
\begin{itemize}
\item \emph{Cleaner}: Removed (i) all heading and trailing text embedded in square or rounded brackets and (ii) heading double character token ending with bracket or fullstop.
\begin{itemize}
\item[(i)]
\begin{Verbatim}
^(?\s?\w{1,5}\s*[):.]\s*
\end{Verbatim}
\item[(ii)] 
\begin{Verbatim}
[\[\(].{1,}[\]\)]
\end{Verbatim}
\end{itemize}
\item \emph{Cleanest}: Accepts only source lines without any punctuation.
\end{itemize}
\end{minipage}
\\ \\
\noindent The original version of the ODIN data contains XML files for 1275 languages, while the cleaner version of ODIN contains IGTs for 1042 languages and the cleanest version contains IGTs for 402 languages. The drop from 1275 to 1042 languages was largely because {\color{red} XXXX} XML files from the original ODIN data had used language codes that were not in ISO 693-3 and for {\color{red} XXXX} other files, the \texttt{<igt>...</igt>} tags were missing.

@@

The Omniglot crawling and cleaning required more tact and after manual inspection of the www.omniglot.com website, only the \emph{`Useful foreign phrases'} and the \emph{`Tower of Babel'} pages were consistent in their HTML markups that allows easy cleaning. We wrote a crawler script that downloads pages with that contains the following in their URLs:

@@

\noindent The primary problem of the Omniglot data is that to the only information about the language of the data on the page is given in the title of the HTML and its URL. Additionally, only the name of the language was given not the ISO 639-3 code. Even though Omniglot pages contain language meta data in XHTML markup attributes, most pages are encoded with the \texttt{lang="en"} attribute because the pages are catered for English speaking audience and the boilerplates (e.g. navigation bar, site banner and footer) on the pages are in English. This causes a problem in file storage and access since it is necessary to save the files in the naming convention, e.g. \texttt{omniglot-smo.txt} but the language code is not found any where on the page. To resolve the problem of standardizing, we have written a conversion script using the code to language mappings from \url{http://www-01.sil.org/iso639-3/iso-639-3.tab}

@@

Although there was a pre-compiled version of the UDHR data from the Natural Language ToolKit (NLTK) corpora distribution\footnote{http://nltk.googlecode.com/svn/trunk/nltk\_data/index.xml}, the distribution was laden with encoding problems during their conversion from pdf to plaintext format. Originally, we tried to resolve the encoding problems by automatically identifying the encoding and decoding the textfiles using the \texttt{libmagic} library\footnote{http://linux.die.net/man/3/libmagic} and then re-encoding them to \texttt{UTF-8} whenever possible.

However the better solution was found when we chanced upon the UDHR site\footnote{http://unicode.org/udhr/d/} from the \url{unicode.org} domain. The UDHR textfiles found on the site was free of encoding problem but there was a boilerplate at the start of each file which records it metadata. It was simply clean by skipping the first four lines of each file.

Although the redistribution of previous compiled data is not scientifically attractive but redistribution/recompilation of data to improve the usability and quality of the data is worth mentioning. For instance, the SETimes corpus first compiled by Tyers and Alperen (2010) was cleaned and redistributed by \newcite{ljubesic2013ner}. [This is not mentioned in the reference...]

@@

Tapping on the crowd-sourced Wikipedia articles for NLP forms a crucial part of developing data-driven NLP tools and application ({\color{red}{citation neeeded}}). To automatically download the Wikipedia dumps, we have scripted the \texttt{wget} with ISO 639-2 language code as a variable (\texttt{\$I}) in a shell script, as such:

@@

After converting them into textfiles, we used the following regexes to delete residual Wikipedia markups and magic words\footnote{http://en.wikipedia.org/wiki/Help:Magic\_words}.

>>>>

Subsequently, \newcite{brunello2009free} stressed the notion of building free corpora from the web using documents released under Creative Commons licenses. To build better quality WaCky corpora, \newcite{versley2012quality} introduced the notion of content-sensitive boilerplate detection in cleaning web corpora.

The initial wave of monolingual WaCky corpora \cite{baroni2009wacky} inspired more domain specific WaCky corpora, such as the Korean Web Learner's Corpus \cite{dickinson2010korean} and the Feed Corpus \cite{minocha2013feed}. Meanwhile, Web-driven corpora compilations also focused on building large  multilingual resources, e.g. The Leipzig Corpora Collection (\texttt{LCC}) \cite{biemann2007leipzig}, and the COrpora from the Web (\texttt{COW}) project \cite{schaefer2012cow}.

Unlike previous WaCs built under the Web as Corpus kool ynitiative (WaCky), we did not perform seed query web-searching or web crawling to achieve a balanced resource. Instead, the foundation text for uniWaC was built using dedicated site crawling and/or site-dependent cleaning from four data sources, viz. (i) \emph{Online Database of Interlinear Text} (ODIN), (ii) \emph{Omniglot} website, (iii) \emph{Universal Declaration of Human Rights} (UDHR) and (iv) \emph{Wikipedia} dumps. Our emphasis for the uniWaC is based on based on \textbf{\emph{universality}}; covering as many languages as possible is the first priority.

@@

\newcite{abney2011data} describe the data structure they envisage for the universal corpus in more detail, distinguishing between \textbf{aligned texts} and \textbf{analysed texts}: the former consists of multiple parallel texts, aligned at the document, sentence, or word level; the latter contains more detailed annotations including parts of speech, morphological information, and syntactic relations. In our work, we have strived to maintain a data structure consistent with their recommendations for aligned texts.

@@

The resulting files were saved as plaintext file with \texttt{UTF-8} encoding. We used one directory for each data source and one file per language. The parallel data (from ODIN and Omniglot) are saved as tab-delimited files, one line per sentence, one column per language. The monolingual data (from UDHR and Wikipedia) are saved as newline delimited plaintext file, each line denoting a paragraph or sentence and each document separated by an empty line. The files are saved with the following naming convention to allow easy access, the filename contains the data source and its ISO 639-3 code, e.g. \texttt{odin-iii.txt} contains the \emph{ODIN} data for the \emph{Nuosu} language. [File structure is not relevant.]

@@

ODIN (The Online Database of Interlinear Text) is a repository of IGTs extracted from scholarly documents \cite{lewis2006odin,lewis2010odin}.  Compared to other resources, it is notable for the breadth of languages included, and the level of linguistic annotation; however, the data requires further processing to bring it in line with the proposed format for a universal corpus.

The ODIN data is easily accessible in XML format from the online database\footnote{http://odin.linguistlist.org/download}, where data for each language is saved in a separate XML file and the Interlinear Glossed Texts (IGTs) are encoded in the \texttt{<igt><example>...</example></igt>} tags. Each XML file is saved under a filename that matches the respective language code. While cleaning the data, filenames that does not adhere to ISO 639-3\footnote{http://www-01.sil.org/iso639-3} were excluded in the uniWaC compilation. [Huh? Should this really happen? - the codes are given on their website and look valid to me...]

@@

a. \quad o lesu mai \\
\indent \qquad 2sg return here \\

\indent \qquad `\emph{You return here.}' \\
\centerline{Figure 1: IGT from ODIN.}
\end{minipage}

\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>21 a. o lesu mai</line>
    <line>2sg return here</line>
    <line>`You return here.'</line>
  </example>
</igit>
\end{lstlisting} 
\centerline{Figure 2: Fijian IGT from ODIN.}
\end{minipage}
\\ \\
\noindent The IGTs from a web document usually follow the Leipzig Glossing Rules, where the first line is the source language text, the second line contains the word/morphemic equivalence and the third line is an English gloss. From the ODIN XML format, an IGT as of Figure 1 will be represented as in an XML snippet as in Figure 2. Eventually, we need to clean and extract (i) the source text `\emph{o lesu mai}' without the preceding index `\emph{a.}' and (ii) the target language gloss without the quotation marks `\emph{You return here}'.
\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>(69) na-Na-tmi-kwalca-t 
    Yimas (Foley 1991)</line>
    <line>3sgA-1sgO-say-rise-PERF
    </line>
    <line>`She woke me up' 
    (by verbal action)</line>
  </example>
</igit>
\end{lstlisting} 
\smallskip
\centerline{Figure 3: Yimas IGT from ODIN.}
\end{minipage}
\\ \\
\noindent The primary problem in extracting the source text is a lack of consistency in the IGTs. In the above examples, the sentence is introduced by a letter or number, which needs to be removed; however, the form of such indexing elements varies. In addition, the source line in Figure 3 includes the language name, and a citation for the example, which introduces noise into the corpus.

Lewis and Xia (2010) noted that it is not uncommon for IGTs found in documents to deviate from the standard convention; when compiling the ODIN data, they attempted a regex-based approached and reported 59\% F-score. Similar to their regex approaches for extracting IGT, we cleaned the source line with regexes and kept the ODIN data with two levels of `cleanliness': [Have we considered that the number of elements in the source line and gloss line should match...?]
\\ \\
\begin{minipage}{\columnwidth}
\begin{itemize}
\item \emph{Cleaner}: Removed (i) all heading and trailing text embedded in square or rounded brackets and (ii) heading double character token ending with bracket or fullstop.
\begin{itemize}
\item[(i)]
\begin{Verbatim}
^(?\s?\w{1,5}\s*[):.]\s*
\end{Verbatim}
\item[(ii)] 
\begin{Verbatim}
[\[\(].{1,}[\]\)]
\end{Verbatim}
\end{itemize}
\item \emph{Cleanest}: Only source lines without punctuation.
\end{itemize}
\end{minipage}
\\ \\

\noindent The original version of the ODIN data contains XML files for 1275 languages, while the cleaner version of ODIN contains IGTs for 1042 languages and the cleanest version contains IGTs for 402 languages. The drop from 1275 to 1042 languages was largely because {\color{red} XXXX} XML files from the original ODIN data had used language codes that were not in ISO 639-3 and for {\color{red} XXXX} other files, the \texttt{<igt>...</igt>} tags were missing.

@@

The Omniglot website\footnote{htp://www.omniglot.com/} is an online encyclopedia of writing systems and languages. We chose to extract information from pages on \emph{`Useful foreign phrases'} and the \emph{`Tower of Babel'} story, both of which give us a parallel data in a reasonably large number of languages.

[The following discussion is far too technical.]

@@

\noindent One problem with this data is that only the language name is given, not the ISO 639-3 code. To resolve this issue, we converted language names to codes using information from the SIL website.\footnote{http://www-01.sil.org/iso639-3/iso-639-3.tab}

@@

The Universal Declaration of Human Rights (UDHR) is a document released by the United Nations, which has been translated into a wide variety of languages. Although there was a pre-compiled version of the UDHR data from the Natural Language ToolKit (NLTK) corpora distribution\footnote{http://nltk.googlecode.com/svn/trunk/nltk\_data/index.xml}, the distribution was laden with encoding problems during their conversion from pdf to plaintext format. Instead, we used the plaintext files available from the Unicode website\footnote{http://unicode.org/udhr/d/}, which are free of encoding issues. The first four lines of each file records metadata, and the rest is the translation of the UDHR.

[This paragraph seems completely incongruous] Although the redistribution of previous compiled data is not scientifically attractive but redistribution/recompilation of data to improve the usability and quality of the data is worth mentioning. For instance, the SETimes corpus first compiled by Tyers and Alperen (2010) was cleaned and redistributed by \newcite{ljubesic2013ner}. [This is not mentioned in the reference...]

@@

Tapping on the crowd-sourced Wikipedia articles for NLP forms a crucial part of developing data-driven NLP tools and application ({\color{red}{citation neeeded}}). [It might be popular, but it's not crucial...] To automatically download the Wikipedia dumps, we have scripted the \texttt{wget} with ISO 639-2 language code as a variable (\texttt{\$I}) in a shell script, as such: [This is too technical]

@@

After converting them into textfiles, we used the following regexes to delete residual Wikipedia markup and so-called ``magic words"\footnote{http://en.wikipedia.org/wiki/Help:Magic\_words}.
